---
title: "Data Book"
output:
  html_document: 
    keep_md: yes
  html_notebook: default
---

```{r setup, include=F, error=T, message=F, warning=F}
knitr::opts_chunk$set(echo=T, comment="#>:", collapse = TRUE,
                      error=T, warning=F, message = F, fig.align = 'center' 
                      )
```



```{r}
# tibble allows to compress the output of a dataset when printing
library(tibble)

dataset <- read.csv("2007_welltest.csv", stringsAsFactors = FALSE)
as.tibble(dataset)   # small output
```


```{r}
# show all the columns 
names(dataset)
```

```{r}
# get only the numeric columns
nums <- sapply(dataset, is.numeric)
numdf <- dataset[, nums]
names(numdf)
```


```{r}
# get only the non-numeric columns
nums <- sapply(dataset, is.numeric)
nonumdf <- dataset[, !nums]
names(nonumdf)
```

```{r}
# show only the ID descriptive variables
library(dplyr)

dataset %>%
    select(CODE, WELL_HOLE_CODE, OP_SUB_AREA_CODE, OP_AREA_CODE, OP_FCTY_1_CODE) %>%
    head
```


```{r}
# the physical well is called a HOLE
# a HOLE can have two strings: LS and SS
# a nominal well will be ANY-A0001-LS and another
# well ANY-A0001-SS. Both belong to a HOLE
# 
# How many holes in this dataset?
length(unique(dataset$WELL_HOLE_CODE))
# unique(dataset$WELL_HOLE_CODE)
```

```{r}
# how many nominal wells
length(unique(dataset$CODE))
```

```{r}
# how many fields
length(unique(dataset$OP_SUB_AREA_CODE))
unique(dataset$OP_SUB_AREA_CODE)
```

```{r}
# how many operating units
length(unique(dataset$OP_AREA_CODE))
unique(dataset$OP_AREA_CODE)
```

```{r}
# how many platforms
length(unique(dataset$OP_FCTY_1_CODE))
unique(dataset$OP_FCTY_1_CODE)
```

```{r}
dataset <- read.csv("2007_welltest.csv", stringsAsFactors = FALSE)
```


`PRODUCTION_DAY`: date that the measurement was taken.

`CODE`: nominal well. In the case of dual wells, it is one of the strings, teh short or the long.

`WELL_HOLE_CODE`: the physical well that is drilled. A hole or physical well could have two nominal wells. See `CODE`.

`WELL_TYPE`:   

        Gas Producer
        Oil Producer
        Gas Producer Cond
        Water Injector
```{r}
unique(dataset$WELL_TYPE)
```

`PROD_METHOD`: How the well produces.
    Natural Flow
    Gas Lift
    Pumping Unit
    
```{r}
unique(dataset$PROD_METHOD)
```


`WELL_CLASS`: Producer or Injector.

```{r}
unique(dataset$WELL_CLASS)
```

`DURATION_HRS`: duration of the test

```{r}
dur <- sort(unique(round(dataset$DURATION_HRS, 0))) # rounding, unique, sort
dur
```


```{r}
# interval analysis of test durantion
table(cut(dataset$DURATION_HRS, breaks = c(0, 1, 2, 4, 6, 12, 24, 48, 72)))
```


`STATUS`: What happens with the well test.
```{r}
unique(dataset$STATUS)
```

`NET_OIL_RATE_ADJ_BBLSPERDAY`: Net oil produced in STB/DAY

```{r}
summary(dataset$NET_OIL_RATE_ADJ_BBLSPERDAY)

# which wells are higher than 10,000 BOPD
lmask <- dataset$NET_OIL_RATE_ADJ_BBLSPERDAY > 10000
imask <- which(lmask)

value = dataset$NET_OIL_RATE_ADJ_BBLSPERDAY[imask]
well  = dataset$CODE[imask]

data.frame(index = imask, well = well, NetOil = value)

```

```{r}
# remove the two highest reading. They belong to a tank battery
netoil <- dataset$NET_OIL_RATE_ADJ_BBLSPERDAY[-c(4579,4959)]
summary(netoil)
```


```{r}
# interval analysis of oil production
table(cut(dataset$NET_OIL_RATE_ADJ_BBLSPERDAY, 
          breaks = c(0, 50, 100, 250, 500, 1000, 2000, 4000, 8000, 10000)))
```

```{r}
# histogram of net oil
hist.data = hist(netoil, plot=F)   # do not plot. 
# log-transform the counts: 2^2, 2^4, ... 2^10 = 1024
hist.data$counts[hist.data$counts>0] <- log(hist.data$counts[hist.data$counts>0], 2)
plot(hist.data)
```

WATERCUT_PCT
```{r}
summary(dataset$WATERCUT_PCT)
```

```{r}
# values that are above 100 watercut
mask  <- dataset$WATERCUT_PCT > 100
value <- dataset$WATERCUT_PCT[which(dataset$WATERCUT_PCT > 100)]

# wells and watercuts that abnormal
well <- dataset$CODE[which(dataset$WATERCUT_PCT > 100)]
data.frame(well = well, watercut = value)
```


```{r}
library(tibble)
# what wells have NAs 
lmask <- is.na(dataset$WATERCUT_PCT)
imask <- which(lmask)

value <- dataset$WATERCUT_PCT[imask]
well  <- dataset$CODE[imask]

as.tibble(data.frame(well = well, watercut = value))
```


```{r}
# interval analysis of watercut
table(cut(dataset$WATERCUT_PCT, 
          breaks = c(0, 10, 25, 50, 80, 90, 95, 100)))

# histogram of only those wells that have watercut < 100
hist(dataset$WATERCUT_PCT[which(!dataset$WATERCUT_PCT > 100)])
```


#### Logical process status of a well
These are character-boolean variables: "Y" or "N".
```{r}
unique(dataset$CAN_PROC_STEAMINJ)
unique(dataset$CAN_PROC_WATINJ)
```


     [15] "CALC_SEQ_NO"                   "CAN_PROC_OIL"                 
     [17] "CAN_PROC_GAS"                  "CAN_PROC_WAT"                 
     [19] "CAN_PROC_GASLIFT"              "CAN_PROC_CO2"                 
     [21] "CAN_PROC_GASINJ"               "CAN_PROC_WATINJ"              
     [23] "CAN_PROC_STEAMINJ"             "CAN_PROC_COND"                
     [25] "CAN_PROC_DILUENT"



#### Logical production status of a well
These are character-boolean variables: "Y" or "N".

     [27] "ISPRODUCER"                    "ISPRODUCEROROTHER"            
     [29] "ISINJECTOR"                    "ISOTHER"                      
     [31] "ISNOTOTHER"                    "ISOILPRODUCER"                
     [33] "ISGASPRODUCER"                 "ISCONDENSATEPRODUCER"         
     [35] "ISWATERPRODUCER"               "ISGASINJECTOR"                
     [37] "ISWATERINJECTOR"               "ISAIRINJECTOR"                
     [39] "ISSTEAMINJECTOR"               "ISWASTEINJECTOR"              
     [41] "ISCO2INJECTOR" 

```{r}
unique(dataset$ISINJECTOR)
unique(dataset$ISGASINJECTOR)
unique(dataset$ISPRODUCEROROTHER)
unique(dataset$ISWASTEINJECTOR)
```

